Conference Paper Title*\\ {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and should not be used} \thanks{Identify applicable funding agency here. If none, delete this.}

### 2023-05-14 LARS_paper/paper.tex
add conclusion

Table \ref{tab:vel_stab} summarizes the collected results. The robot did not reach a velocity of 0.05 m/s in any test. This result is due to errors in the final position of each leg during the swing phase: the legs could not reach the desired $(x,y)$ position required to maintain the desired velocity. The fact that the final position of one leg is a function of its motors' final orientation, indicates that the motors did not reach their desired orientation. Prior to the experiment, the gait movement was tested with the legs off the ground (almost no load), and no significant errors were noted. Since the errors only appeared while walking, it is perhaps caused by the fact that the motors were not powerful enough to support the robot's weight appropriately while walking.

@@

\section{Conclusão}

Este trabalho abordou os conceitos utilizados para locomoção de robôs quadrúpedes e buscou aplicá-los no desenvolvimento de um robô real. O Caramelo foi desenvolvido para fins de educação e pesquisa na área de robôs com pernas, mais especificamente quadrúpedes. Trata-se de um projeto \textit{open source}, cujo código fonte está disponível publicamente no \textit{GitHub} \cite{CaramelRepo}.

Os testes e experimentos realizados buscaram avaliar a performance da locomoção do robô no espaço tridimensional. Os testes preliminares indicaram que o robô possui a capacidade de mover o corpo em 6 GDL e controlar sua orientação em $roll$ e $pitch$. Além disso, pôde-se concluir que a pata é capaz de seguir uma trajetória até um ponto requisitado em $x$ e $y$ no cenário sem carga. Não foram constatadas diferenças significativas no tempo total e na altura máxima da trajetória nos dois testes realizados. Contudo, foi observado um atraso no tempo de execução trajetória, relacionado ao atraso na resposta dos motores.

Os experimentos de locomoção mostraram que o robô não alcançou a velocidade desejada em nenhum dos testes, estando  a velocidade média do teste no terreno plano com o controle de angulação ativo a mais próxima deste valor. Essa diferença pode estar relacionada com o fato de que o sistema de controle não conseguiu operar de forma adequada quando sustentando o peso do robô. Como o controle da trajetória é feito em malha aberta (apenas os controladores individuais das juntas possuem malha fechada), não há compensação, e a pata não alcança a distância esperada. Ademais, também pôde-se observar que o controle de angulação resultou numa maior estabilidade do robô em ambos os terrenos.

\section{Considerações Finais}

Para trabalhos futuros, recomenda-se a revisão dos atuadores usados no robô. Ao longo da fase de testes, foi observado que estes frequentemente falhavam por conta de sobrecarga, o que é um sinal de que foram subdimensionados. Por consequência, a capacidade de \textit{payload} não é significativa. Uma possível solução para essa questão é a substituição do modelo atual por um com maior torque e/ou a investigação da configuração das pernas do robô. A equipe observou que, como os motores que mais falhavam eram os traseiros, adotar a configuração \textit{elbow-knee} pode diminuir a carga neles, já que deixaria as pernas traseira em uma configuração simétrica às frontais. Espera-se, também, que isso influencie na performance de locomoção, deixando o robô mais rápido e estável. Além disso, sugere-se a aplicação de outros métodos de controle, como controladores baseados em modelos dinâmicos ou até em \textit{machine learning}.

>>>>

Table \ref{tab:vel_stab} summarizes the collected results. The robot did not reach a velocity of 0.05 m/s in any test. This result is due to errors in the final position of each leg during the swing phase: the legs could not reach the desired $(x,y)$ position required to maintain the desired velocity. The fact that the final position of one foot is a function of its motors' final orientation, indicates that the motors did not reach their desired orientation. Prior to the experiment, the gait movement was tested with the legs off the ground (almost no load), and no significant errors were noted. Since the errors only appeared while walking, it is perhaps caused by the fact that the motors were not powerful enough to support the robot's weight appropriately while walking.

@@

\section{Conclusion}

This paper addressed the design, prototyping and testing of a quadruped robot. An experiment was held to assess the robot's walking performance. The results showed that the robot did not reach the desired velocity in any of the tests. This was due to errors during the gait's swing phase, likely caused by the external disturbances caused by the robot's weight. This disturbances were not compensated, because the gait trajectory traking is open-loop (only the indivual joint's controllers are closed-loop). In addition to that, it was passible to conclude that the implemented stabilization controller was able to lower the robot's body oscillations while walking in pitch and roll in both flat and irregular terrains.

Reviewing the robot's actuators is recommended for future work. Perhaps a model with more torque is required for a more precise control performance.

The robot's project is open source and is available on GitHub \cite{CaramelRepo}.
